---
title: "Using plotly for animation and more"
author: "Robert Carver"
output:
  html_document:
    df_print: paged
---

# Introduction
In the most recent DataCamp lesson, you worked with a so-called `ggplot2` *extension* called `gganimate` to introduce movement into a ggplot graph. It turns out that several other programs and packages are necessary for gganimate. An alternative is the package `plotly` which "wraps" around `ggplot2` objects and allows for interactivity and animation with techniques like brushing, zooming, panning, animation and more.

## This document
This page was built as an RNotebook using `rmarkdown` and then "knitted" with `knitr` . Most of the examples and code are taken and adapted from *The Plotly Cookbook* by Carson Sievert, available online at https://plotly-book.cpsievert.me/index.html


# The code and examples
We start by installing and calling several r packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(plotly)
library(dplyr)
library(gapminder)
```

The first few examples use some data about Texas housing prices, which is bundled with `ggplot2`. The relevant data frame is called `txhousing`.

1.  Several ways to manage overplotting, with brushing added:


```{r, warning =FALSE, message=FALSE, error=FALSE}
p <- ggplot(txhousing, aes(date, median)) +
     geom_line(aes(group = city), alpha = 0.2)
subplot(
     p, ggplotly(p, tooltip = "city"), 
     ggplot(txhousing, aes(date, median)) + geom_bin2d(),
     ggplot(txhousing, aes(date, median)) + geom_hex(),
     nrows = 2, shareX = TRUE, shareY = TRUE,
     titleY = FALSE, titleX = FALSE
)
```

2.  There are many cities in Texas. Here is a multi-line graph, highlighting one city.

```{R}

tx <- group_by(txhousing, city)
# initiate a plotly object with date on x and median on y
p <- plot_ly(tx, x = ~date, y = ~median)


# add a line highlighting houston
add_lines(
     # plots one line per city since p knows city is a grouping variable
     add_lines(p, alpha = 0.2, name = "Texan Cities", hoverinfo = "none"),
     name = "Houston", data = filter(txhousing, city == "Houston")
)

```

### More examples using the mtcars data frame

You may recall the `mtcars` data frame in r. It contains data about car characteristics, compiled by Motor Trend magazine.

```{r}
glimpse(mtcars)
head(mtcars)
```

1.  Allow panning (moving the entire plot within a window)

```{r, message=FALSE, error=FALSE}
p <- ggplot(fortify(mtcars), aes(wt, mpg)) + 
     geom_point() + geom_smooth()
gg <- ggplotly(p)
layout(gg, dragmode = "pan")

```


2.  Add a range slider, as in Google Finance


```{r, message=FALSE, error=FALSE}
p <- ggplot(fortify(mtcars), aes(wt, mpg)) + 
     geom_point() + geom_smooth()
gg <- ggplotly(p)
rangeslider(gg)
```


3.  How about a 3-D scatterplot?

```{r}
plot_ly(mpg, x = ~cty, y = ~hwy, z = ~cyl) %>%
  add_markers(color = ~cyl)
```  
  
4.  More animation, a la Tableau and Hans Rosling

```{r, message=FALSE, error=FALSE}
data(gapminder, package = "gapminder")

meanLife <- with(gapminder, tapply(lifeExp, INDEX = continent, mean))
gapminder$continent <- factor(
     gapminder$continent, levels = names(sort(meanLife))
)
base <- gapminder %>%
     plot_ly(x = ~gdpPercap, y = ~lifeExp, size = ~pop, 
             text = ~country, hoverinfo = "text") %>%
     layout(xaxis = list(type = "log"))

base %>%
     add_markers(color = ~continent, alpha = 0.2, showlegend = F) %>%
     add_markers(color = ~continent, frame = ~year, ids = ~country) %>%
     animation_opts(1500, redraw = FALSE)
```
   
#  Conclusion

Animation and interactivity can help in exploration, presentation, and interactive story-telling. Several R pacakges are available to facilitate the process, and in some cases to offer capabilities not yet available in Tableau.
